Optimal (Multiway) Spatial Joins

### 2024-05-24 msj.tex
yufei

\begin{figure*}
    \begin{tabular}{cccc}
        \includegraphics[height=30mm]{./artwork/prob-a} & 
        \hspace{3mm}
        \includegraphics[height=27mm]{./artwork/prob-b} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-c} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-d} \\[2mm]
        (a) Problems $\EuScript{A}$ &
        \hspace{3mm}
        (b) Problem $\EuScript{B}$ &
        \hspace{3mm}
        (c) Problem $\EuScript{C}$
        &
        \hspace{3mm}
        (d) Problem $\EuScript{D}$
    \end{tabular}

@@

\extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $P$ of points and set $R$ of rectangles, all in 2D space. In the {\em detection version} of Problem $\EuScript{A}$, the goal is to output, for each point $p \in P$, whether it is covered by at least one rectangle in $R$. Figure~\ref{fig:probs}a gives an example where $P = \set{p_1, p_2, p_3}$ and $R = \set{r_1, r_2}$; the output is ``yes'' for $p_2$ and $p_3$ and ``no'' for $p_1$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time, where $n = |P| + |R|$.

@@

\extraspacing {\bf Problem $\bm{\EuScript{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments, all in 2D space. The goal is to report, for each segment $h \in R$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$. If $h$ does not intersect with any segment in $V$, report nothing for $h$. Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |V|$.

\extraspacing {\bf Problem $\bm{\EuScript{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the left end point of $h$ and the point $(x, y)$. If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$. Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.

@@

In Appendix~\ref{app:prob-b}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.

\extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$ --- formally, $h = [x_1, x_2] \times y$ has the smallest $y$ coordinate among all the segments in $H$ crossing $r$. If such an $h$ does not exist, output nothing for $r$.  Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.

In Appendix~\ref{app:prob-a}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.

\extraspacing {\bf Problem $\bm{\EuScript{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, all the rectangles $r \in R$ containing $h$. Furthermore, those rectangles need to be sorted by the x-coordinates of their right boundaries. Formally, if $r_1, r_2, ..., r_t$ (for some $t \ge 1$) are all the rectangles in $R$ containing $h$, we output $(h: r_1, r_2, ..., r_t)$, provided that $x_i > x_{i-1}$ for each $i \in [2, t]$ where $x_i$ (resp., $x_{i-1}$) is the x-coordinate of the right boundary of $r_i$ (resp., $r_{i-1}$). Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$.

In Appendix~\ref{app:prob-c}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.

\extraspacing {\bf Problem $\bm{\EuScript{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each rectangle $r \in R$, all the segments $h \in H$ crossing $r$. Furthermore, those segments need to be sorted by their y-coordinates. Formally, if $h_1, h_2, ..., h_t$ (for some $t \ge 1$) are all the segments in $H$ crossing $r$, we output $(r: h_1, h_2, ..., h_t)$, provided that $y_i > y_{i-1}$ for each $i \in [2, t]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$). See Figure~\ref{fig:probs}a again where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1: h_3, h_2), (r_2: h_2, h_1)}$.

In Appendix~\ref{app:prob-d}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $h$ crosses $r$.

@@

    \item $v_1$ (resp., $v_2$): the leftmost (resp., rightmost) segment --- i.e., with the smallest (resp., largest) x-coordinate --- among all the segments in $V$ that cross $r$.
}
Define $r_1'$ as the rectangle defined by the above four segments in the way shown in Figure~\ref{fig:hv:type1}(a) ($r_1'$ is the shaded area). Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$. Define
\myeqn{
    R_1' &=& \set{r_1' \mid r_1 \in R_1 \text{ and its $r_1'$ defined as above exists}}. \nn
}
Computing the corresponding ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\EuScript{A}$. By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$. It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time.

@@

\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{A}}$}

\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{B}}$}

\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{C}}$}

\extraspacing {\bf Algorithm for Problem $\bm{\EuScript{D}}$}

>>>>

\begin{figure*}
    \begin{tabular}{ccccc}
        \includegraphics[height=30mm]{./artwork/prob-a} & 
        \hspace{3mm}
        \includegraphics[height=27mm]{./artwork/prob-b} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-c} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-d} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-e} \\[2mm]
        (a) Problems $\mathscr{A}$ &
        \hspace{3mm}
        (b) Problem $\mathscr{B}$ &
        \hspace{3mm}
        (c) Problem $\mathscr{C}$ &
        \hspace{3mm}
        (d) Problem $\mathscr{D}$ &
        \hspace{3mm}
        (3) Problem $\mathscr{E}$
    \end{tabular}

@@

\extraspacing {\bf Problem $\bm{\mathscr{A}}$.} The input involves a set $P$ of points and set $R$ of rectangles, all in 2D space. In the {\em detection version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, whether it is covered by at least one rectangle in $R$. Figure~\ref{fig:probs}a gives an example where $P = \set{p_1, p_2, p_3}$ and $R = \set{r_1, r_2}$; the output is ``yes'' for $p_2$ and $p_3$ and ``no'' for $p_1$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time, where $n = |P| + |R|$.

@@

\extraspacing {\bf Problem $\bm{\mathscr{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments, all in 2D space. The goal is to report, for each segment $h \in R$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$. If $h$ does not intersect with any segment in $V$, report nothing for $h$. Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |V|$.

\extraspacing {\bf Problem $\bm{\mathscr{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the point $(x_1, y)$ and the point $(x, y)$. If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$. Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.

\extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. In the {\em first-crossing} version of the problem, the goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$ --- formally, $h = [x_1, x_2] \times y$ has the smallest $y$ coordinate among all the segments in $H$ crossing $r$. If such an $h$ does not exist, output nothing for $r$.  Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.

@@

In the {\em all-crossing} version of the problem, the goal is to report, for each rectangle $r \in R$, all the segments $h \in H$ crossing $r$. Furthermore, those segments need to be sorted by their y-coordinates. Formally, if $h_1, h_2, ..., h_z$ (for some $z \ge 1$) are all the segments in $H$ crossing $r$, we output $(r: h_1, h_2, ..., h_z)$, provided that $y_i \ge y_{i-1}$ for each $i \in [2, z]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$). In the example of Figure~\ref{fig:probs}d, the output is $\set{(r_1: h_3, h_2), (r_2: h_2, h_1)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $\out$ is the number of pairs $(h, r) \in H \times R$ such that $h$ crosses $r$.

\extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, all the rectangles $r \in R$ containing $h$. Furthermore, those rectangles need to be sorted by the x-coordinates of their right boundaries. Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $R$ containing $h$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$. Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.

@@

    \item $v_1$ (resp., $v_2$): the leftmost (resp., rightmost) segment --- i.e., with the smallest (resp., largest) x-coordinate --- among all the segments in $V$ that cross $r$.
}
Define $r_1'$ as the rectangle defined by the above four segments in the way shown in Figure~\ref{fig:hv:type1}(a) ($r_1'$ is the shaded area). Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$. Define
\myeqn{
    R_1' &=& \set{r_1' \mid r_1 \in R_1 \text{ and its $r_1'$ defined as above exists}}. \nn
}
Computing the corresponding ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\mathscr{A}$. By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$. It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time.

@@

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{A}}$}

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{B}}$}

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{C}}$}

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{D}}$}

\extraspacing {\bf Algorithm for Problem $\bm{\mathscr{E}}$}
